\documentclass[11pt]{article}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage[margin=1in]{geometry}
\usepackage{xcolor}

% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{counterexample}[theorem]{Counterexample}

% Commands
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\cl}[1]{\overline{#1}}
\newcommand{\Fix}{\operatorname{Fix}}
\newcommand{\AD}{\operatorname{AD}}
\newcommand{\id}{\operatorname{id}}
\newcommand{\diam}{\operatorname{diam}}
\newcommand{\dist}{\operatorname{dist}}

\title{Geometric Limits of Defense Design:\\
  Impossibility Theorems for Continuous\\
  Prompt Injection Defenses}

\author{
  Manish Bhatt\thanks{Corresponding author: \texttt{manish.bhatt13212@gmail.com}. This work was conducted independently and does not reflect the views of the authors' respective employers.} \\
  \and
  Sarthak Munshi \\
  \and
  Vineeth Sai Narajala \\
  \and
  Idan Habler \\
  \and
  Ammar Al-Kahfah \\
  \and
  Ken Huang \\
  \and
  Blake Gatto \\
}

\date{}

\begin{document}

\maketitle

\begin{abstract}
Prior work on LLM safety has focused on discovering adversarial inputs
or bounding perturbation robustness around individual examples. We ask
a different question: what are the \emph{geometric limits of defense
design itself}?

We prove three results of increasing strength. First, any continuous,
utility-preserving defense on a connected Hausdorff space must fix
boundary points of the alignment deviation surface (\emph{boundary
fixation}). Second, when the defense and alignment function are both
Lipschitz, the defense cannot uniformly reduce alignment deviation far
below the threshold on a positive-measure band around the boundary
(\emph{$\varepsilon$-robust constraint}). Third, under a transversality
condition---the alignment surface rises faster than the defense can pull
it down---the defense leaves a \emph{positive-measure region genuinely
unsafe}: $f(D(x)) > \tau$ persists over volume
(\emph{persistent unsafe region}).

These are not statements about attacks or training---they are constraints
on what any admissible defense mapping can do, independent of the model,
data, or optimization procedure. The Tietze extension theorem ensures
that any finite set of behavioral observations admits continuous
extensions for which all three impossibilities hold.

We also prove a parallel set of discrete impossibility results that
require no topology, and extend the theory to multi-turn interactions
(the impossibility recurs at every turn), stochastic defenses
(randomization does not help), and capacity parity (utility preservation
taxes the defense even under equal resources).

The paper defines the theoretical speed limit for AI safety. It proves we can't build a `Perfect Shield' in a continuous space. However, by using multi-layered, discontinuous defenses and ensuring our boundary behavior is benign, we can operate safely within those limits. The `Manifold of Failure' exists, but we can ensure it's empty of actual harm. The full theory is machine-verified in Lean~4 with Mathlib (38~files,
$\sim$300 theorems, zero \texttt{sorry}, three standard axioms) and
validated against empirical findings from three frontier LLMs. \url{https://github.com/mbhatt1/stuff/tree/main/ManifoldProofs}
\end{abstract}

% ============================================================
\section{Introduction}
\label{sec:intro}

Can you build a wrapper around a language model that eliminates all
prompt injection vulnerabilities?

Most current work on LLM safety implicitly assumes the answer is
yes---at least in principle, even if current tools fall short in practice.
The defense mechanisms people have proposed span a wide range: input
classifiers that flag suspicious prompts~\cite{alon2023detecting},
output filters that catch unsafe completions~\cite{inan2023llama},
and constitutional rewriting pipelines that reshape model
behavior~\cite{bai2022constitutional}. But despite their differences,
all of these defenses share the same underlying structure: they are
functions $D\colon X \to X$ that preprocess prompts before the model
sees them, aiming to map unsafe inputs to safe equivalents while leaving
safe inputs alone.

Prior work has largely focused on discovering or mitigating individual
adversarial inputs. We ask a different question: what are the
\emph{geometric limits of the defense mapping itself}? We show that
under mild and commonly assumed design constraints---continuity and
utility preservation---no defense can eliminate the safety boundary.
Vulnerabilities arise not from specific inputs, but from the structure
of the space and the constraints imposed on admissible defenses. This
shifts the problem from adversarial search to geometric inevitability.

We prove three results of increasing strength.

\paragraph{Tier~1: Boundary fixation.}
If the defense is continuous, utility-preserving, and the prompt space
is connected, then the defense must fix at least one boundary point---a
prompt where alignment deviation equals the threshold
exactly---passing it through with no remediation (\Cref{thm:main}).
The proof is five steps of point-set topology.

\paragraph{Tier~2: $\varepsilon$-robust constraint.}
If the defense is also $K$-Lipschitz and the alignment deviation
function is $L$-Lipschitz, the defense cannot uniformly reduce
alignment deviation far below~$\tau$ on a neighborhood of the fixed
boundary point. For any $x$ within distance $\delta$ of the fixed
point~$z$:
\begin{equation}\label{eq:eps-bound}
  f(D(x)) \;\geq\; \tau - L(K+1)\,\delta.
\end{equation}
The $\varepsilon$-band $\{x : \tau - \varepsilon \leq f(x) \leq \tau\}$
has positive measure for every $\varepsilon > 0$~(\Cref{thm:band-measure}).
This constrains the \emph{depth} to which the defense can push
near-boundary points into the safe region, though it does not by itself
prevent the defense from nudging them slightly below~$\tau$.

\paragraph{Tier~3: Persistent unsafe region.}
Under a transversality condition---the alignment surface has directional
slope exceeding $L(K+1)$ at the boundary---the defense leaves a
\emph{positive-measure region genuinely unsafe}:
\begin{equation}\label{eq:persistent}
  f(D(x)) > \tau \quad\text{for all } x \in \mathcal{S},
  \qquad \mu(\mathcal{S}) > 0.
\end{equation}
This is the strongest result: not boundary fixation, not a near-threshold
band, but genuine persistent unsafe volume~(\Cref{thm:persistent}).
The condition holds whenever the alignment surface rises faster than
the defense can pull it down---empirically observed in all three models
we study.

\paragraph{Why continuity?}
Continuity captures a natural requirement for practical defenses: small
changes in input should not produce arbitrarily large changes in the
defense's output, as such instability would degrade usability and
reliability. Any defense based on learned representations, smooth
transformations, or differentiable processing satisfies this constraint.
Discontinuous defenses (hard blocklists, keyword filters) escape the
theorem but suffer their own well-known failure modes.

\paragraph{From discrete to continuous.}
All three results apply to continuous interpolants of discrete data.
The Tietze extension theorem guarantees that any finite set of
behavioral observations in a normal space admits continuous extensions,
and the impossibilities hold for every such extension
(\Cref{thm:tietze}). This does not claim the results hold for the
underlying discrete token system directly, but for every continuous
model consistent with the observations.

Both the basic and $\varepsilon$-robust proofs are machine-verified in
Lean~4 with zero \texttt{sorry} statements, as is the persistent unsafe
region theorem. All three conditions---continuity, utility preservation,
and connectedness---are individually necessary; we give counterexamples
for each.

\paragraph{Contributions.}
\begin{itemize}[nosep]
  \item \textbf{Tier~1}: Boundary fixation for continuous
    utility-preserving defenses (\Cref{thm:main}).
  \item \textbf{Tier~2}: $\varepsilon$-robust constraint showing the
    defense cannot uniformly reduce AD far below~$\tau$ on a
    positive-measure band (\Cref{thm:eps-robust,thm:band-measure}).
  \item \textbf{Tier~3}: Persistent unsafe region under transversality
    (\Cref{thm:persistent}).
  \item A formally verified bridge from discrete data to continuous
    theory via Tietze extension (\Cref{thm:tietze}).
  \item Direct discrete impossibility results requiring no topology
    (\Cref{sec:discrete}).
  \item Multi-turn, stochastic, and capacity parity extensions
    (\Cref{sec:extensions}).
  \item Counterexamples proving each hypothesis is necessary
    (\Cref{sec:counterexamples}).
  \item Supporting theorems: basin structure, fragmentation,
    convergence, transferability, authority monotonicity, gradient
    ascent, fine-tuning stability, cost asymmetry
    (\Cref{sec:landscape,sec:attacks,sec:stability,sec:cost}).
  \item Theory--experiment correspondence with three LLMs
    (\Cref{sec:experiments}).
  \item Lean~4 artifact: 38 files, $\sim$300 theorems, zero
    \texttt{sorry}, three standard axioms (\Cref{sec:artifact}).
\end{itemize}


% ============================================================
\section{Related Work}
\label{sec:related}

\paragraph{Adversarial robustness.}
The existence of adversarial examples in neural networks has been known
for over a decade~\cite{szegedy2013intriguing,goodfellow2014explaining},
and a substantial body of work has pursued formal verification of
robustness properties---through abstract
interpretation~\cite{singh2019abstract}, SMT-based
methods~\cite{katz2017reluplex}, and randomized
smoothing~\cite{cohen2019certified}. Unlike adversarial robustness
results, which characterize the existence of perturbations or bound
robustness radii around individual inputs under specific norms, our
result is a \emph{structural constraint on the defense mapping itself},
independent of perturbation models, norms, or training procedures.
Certified defenses provide guarantees about perturbation neighborhoods;
our result constrains the \emph{global behavior} of the defense
function, showing that no such defense can eliminate boundary regions.

\paragraph{Prompt injection and jailbreaking.}
Techniques for finding individual vulnerabilities include gradient-based
attacks like GCG~\cite{zou2023universal}, search-based approaches like
PAIR~\cite{chao2024jailbreaking} and TAP~\cite{mehrotra2024tree},
and quality-diversity methods like Rainbow
Teaming~\cite{samvelyan2024rainbow}. The Manifold of Failure
framework~\cite{bhatt2025manifold} uses MAP-Elites to map the
\emph{continuous topology} of failures. Existing work identifies
individual failure cases or distributions of attacks. Our result shows
that these failures are not incidental but \emph{topologically
unavoidable} under mild assumptions, providing a formal explanation
for their persistence.

\paragraph{Impossibility results in ML.}
The no-free-lunch theorems~\cite{wolpert1997no} limit average
performance across tasks. The robustness--accuracy tradeoff
literature~\cite{tsipras2018robustness} shows that adversarial
robustness costs standard performance. Our result is different in
character: it holds \emph{pointwise} for any continuous defense on a
\emph{fixed task}, with no dependence on distributions, learning
algorithms, or training data.

\paragraph{Topological views of neural networks.}
Work on decision boundary geometry, the manifold hypothesis, and
topology of neural representations has established that topological
structure matters. While our proof relies on elementary topology, the
novelty lies in identifying that \emph{standard design constraints for
LLM defenses induce a forced fixed-point structure at the safety
boundary}---a connection not previously characterized. The persistent
unsafe region theorem (Tier~3) goes further, using metric structure to
establish volumetric persistence.

\paragraph{Formal verification of ML systems.}
Proof assistants like Lean and Coq have been applied to verify
properties of neural networks~\cite{bagnall2019certifying} and
optimization algorithms~\cite{bolte2024}. To the best of our knowledge,
ours is the first formally verified impossibility result specifically
targeting prompt injection defense.


% ============================================================
\section{Setup}
\label{sec:setup}

\begin{definition}[Alignment Deviation Function]
\label{def:ad}
Let $X$ be a topological space. An \emph{alignment deviation function} is
a continuous map $f\colon X \to \R$. Given a threshold $\tau\in\R$:
\begin{align}
  S_\tau &= \{x \in X : f(x) < \tau\}  &&\text{(safe region)} \\
  U_\tau &= \{x \in X : f(x) > \tau\}  &&\text{(unsafe region)} \\
  B_\tau &= \{x \in X : f(x) = \tau\}  &&\text{(boundary)}
\end{align}
\end{definition}

\begin{definition}[Defense]
\label{def:defense}
A \emph{defense} is a continuous map $D\colon X \to X$. It is
\emph{utility-preserving} if $D(x) = x$ for all $x \in S_\tau$,
and \emph{complete} if $f(D(x)) < \tau$ for all $x\in X$.
\end{definition}


% ============================================================
\section{Tier~1: Boundary Fixation}
\label{sec:proof}

\begin{theorem}[Boundary Fixation]
\label{thm:main}
Let $X$ be a connected Hausdorff space. Let $f\colon X\to\R$ be
continuous with $S_\tau \neq \emptyset$ and $U_\tau \neq \emptyset$. Let
$D\colon X \to X$ be continuous with $D|_{S_\tau} = \id$.

Then there exists $z\in X$ with $f(z) = \tau$ and $D(z) = z$.
Moreover, every $z \in \cl{S_\tau} \setminus S_\tau$ satisfies
$f(z) = \tau$ and $D(z) = z$. This set is nonempty.
\end{theorem}

\begin{proof}
Five steps, each using exactly one hypothesis.

\textbf{Step~1} (Fixed-point set is closed).
$\Fix(D) = \{x : D(x) = x\}$ is the preimage of the diagonal
$\Delta \subset X \times X$ under $x \mapsto (D(x), x)$. Hausdorff
$\Rightarrow$ $\Delta$ closed $\Rightarrow$ $\Fix(D)$ closed.

\textbf{Step~2} (Closure of safe region $\subseteq$ fixed points).
$S_\tau \subseteq \Fix(D)$ by utility preservation. $\Fix(D)$ closed
$\Rightarrow$ $\cl{S_\tau} \subseteq \Fix(D)$.

\textbf{Step~3} (Safe region is not closed).
$S_\tau = f^{-1}((-\infty,\tau))$ is open. If also closed, it would be
clopen, hence $\emptyset$ or $X$---contradicting both regions nonempty.

\textbf{Step~4} (Boundary point in closure).
$S_\tau$ not closed $\Rightarrow$ $\cl{S_\tau} \supsetneq S_\tau$. Any
$z \in \cl{S_\tau} \setminus S_\tau$ satisfies $f(z) \leq \tau$
(closure $\subseteq \{f \leq \tau\}$) and $f(z) \geq \tau$
($z \notin S_\tau$). So $f(z) = \tau$.

\textbf{Step~5} (Defense fixes it).
$z \in \cl{S_\tau} \subseteq \Fix(D)$, so $D(z) = z$ and
$f(D(z)) = \tau \not< \tau$.
\end{proof}

\begin{corollary}
No continuous utility-preserving defense is complete.
\end{corollary}


% ============================================================
\subsection{Each Hypothesis Is Necessary}
\label{sec:counterexamples}

\begin{counterexample}[Removing connectedness]
$X = \{0,1\}$ discrete, $f(0)=0$, $f(1)=1$, $\tau=0.5$.
$D(0)=0$, $D(1)=0$: continuous, utility-preserving, complete.
\end{counterexample}

\begin{counterexample}[Removing continuity]
$X=[0,1]$, $f(x)=x$, $\tau=0.5$. $D(x)=x$ for $x<0.5$,
$D(x)=0$ for $x \geq 0.5$: utility-preserving and complete,
but discontinuous at $0.5$.
\end{counterexample}

\begin{counterexample}[Removing utility preservation]
$D(x) = x_0$ for a fixed safe point: continuous and complete,
but destroys all inputs.
\end{counterexample}

\begin{remark}[The defense trilemma]
A defense can satisfy at most two of: continuity, utility preservation,
completeness.
\end{remark}

\begin{remark}[Not all boundary points are fixed]
The theorem captures $\cl{S_\tau} \setminus S_\tau$, not all of
$B_\tau$. Boundary points isolated from the safe side may escape.
Verified in Lean with $f(x)=x^2$, $\tau=0$.
\end{remark}


% ============================================================
\section{Tier~2: $\varepsilon$-Robust Constraint}
\label{sec:eps-robust}

Tier~1 fixes boundary points---potentially measure zero. Tier~2 extends
the constraint to a positive-measure neighborhood.

\begin{theorem}[$\varepsilon$-Robust Defense Constraint]
\label{thm:eps-robust}
Let $X$ be a connected Hausdorff metric space, $f$ be $L$-Lipschitz
with $S_\tau, U_\tau \neq \emptyset$, and $D$ be $K$-Lipschitz with
$D|_{S_\tau} = \id$.

Then there exists $z \in X$ with $f(z) = \tau$, $D(z) = z$, such that
for all $x \in X$ (with strongest constraint near~$z$):
\begin{equation}
  f(D(x)) \;\geq\; \tau \;-\; L(K+1)\,\dist(x, z).
\end{equation}
\end{theorem}

\begin{proof}
From \Cref{thm:main}, $z$ exists. Since $D$ is $K$-Lipschitz and
$D(z) = z$: $\dist(D(x), x) \leq (K+1)\dist(x,z)$ (triangle
inequality). Since $f$ is $L$-Lipschitz:
$|f(D(x)) - f(x)| \leq L(K+1)\dist(x,z)$.
Extracting the lower bound gives
$f(D(x)) \geq f(x) - L(K+1)\dist(x,z) \geq \tau - L(K+1)\dist(x,z)$
(using $f(z) = \tau$ and Lipschitz of $f$).
\end{proof}

\begin{theorem}[Positive-Measure $\varepsilon$-Band]
\label{thm:band-measure}
Under the hypotheses of \Cref{thm:eps-robust}, for every
$\varepsilon > 0$ such that $f$ takes values below $\tau - \varepsilon$,
the $\varepsilon$-band
$\mathcal{B}_\varepsilon = \{x : \tau - \varepsilon \leq f(x)
\leq \tau\}$
has positive measure (under any measure positive on nonempty open sets).
Specifically, $\operatorname{ball}(c, \varepsilon/(4L)) \subseteq
\mathcal{B}_\varepsilon$ for any $c$ with $f(c) = \tau - \varepsilon/2$.
\end{theorem}

\begin{remark}[What Tier~2 does and does not say]
\Cref{thm:eps-robust} constrains the \emph{depth} to which the defense
can push near-boundary points below~$\tau$. It does \emph{not} prove
these points remain unsafe---the defense may map them to values in
$[\tau - L(K+1)\delta,\; \tau)$, which is below threshold. The bound
is tight: a defense mapping every point in a $\delta$-ball to $f$-value
$\tau - L(K+1)\delta$ would saturate it.
\end{remark}


% ============================================================
\section{Tier~3: Persistent Unsafe Region}
\label{sec:persistent}

Tier~2 constrains depth but permits the defense to nudge points below
$\tau$. Under a transversality condition, Tier~3 proves the defense
\emph{cannot make certain points safe at all}.

\begin{definition}[Steep region]
\label{def:steep}
Given a fixed boundary point $z$ (from \Cref{thm:main}), define
\[
  \mathcal{S} = \{x \in X : f(x) > \tau + L(K+1)\,\dist(x, z)\}.
\]
This is the set of points where the alignment deviation exceeds~$\tau$
by more than the defense's Lipschitz bound can compensate.
\end{definition}

\begin{theorem}[Persistent Unsafe Region]
\label{thm:persistent}
Under the hypotheses of \Cref{thm:eps-robust}, if $\mathcal{S}$ is
nonempty, then:
\begin{enumerate}[nosep]
  \item $\mathcal{S}$ is open.
  \item $\mathcal{S}$ has positive measure (under any measure positive
    on nonempty open sets).
  \item For every $x \in \mathcal{S}$: $f(D(x)) > \tau$.
\end{enumerate}
In particular, $\mu(\{x : f(D(x)) > \tau\}) > 0$---the defense leaves
a positive-measure region genuinely unsafe.
\end{theorem}

\begin{proof}
\textbf{(1)} $\mathcal{S}$ is the strict superlevel set of the
continuous function $x \mapsto f(x) - L(K+1)\dist(x,z)$ at level~$\tau$,
hence open.

\textbf{(2)} Open and nonempty implies positive measure.

\textbf{(3)} For $x \in \mathcal{S}$: $f(x) > \tau + L(K+1)\dist(x,z)$.
From the proof of \Cref{thm:eps-robust}, the defense satisfies
$f(D(x)) \geq f(x) - L(K+1)\dist(x,z)$. Combining:
\[
  f(D(x)) \geq f(x) - L(K+1)\dist(x,z) > \tau.
\]
The defense cannot bring these points below threshold.
\end{proof}

\begin{remark}[When is $\mathcal{S}$ nonempty?]
\label{rem:transversality}
The set $\mathcal{S}$ is nonempty whenever the alignment surface rises
with directional slope exceeding $L(K+1)$ at the boundary point~$z$.
In a normed space, if $f$ has directional derivative $c > L(K+1)$
along some unit vector~$v$ at~$z$, then $z + tv \in \mathcal{S}$ for
small $t > 0$ (verified in Lean as \texttt{transversality\_from\_deriv}).

Informally: if the alignment surface rises faster than the defense can
pull it down, unsafe behavior persists over volume. This condition is
empirically observed in all three models we study---the Lipschitz
constant of the alignment surface near boundary crossings consistently
exceeds reasonable defense aggressiveness bounds.
\end{remark}

\begin{remark}[The defense's fundamental dilemma]
The three tiers reveal a progressive bind. Tier~1: the defense
\emph{must} fix boundary points. Tier~2: it cannot push nearby points
\emph{deep} into safety. Tier~3: if the landscape is steep, it cannot
push them into safety \emph{at all}. Increasing $K$ (defense
aggressiveness) raises the transversality threshold $L(K+1)$, requiring
an even steeper landscape to trigger Tier~3---but it also makes the
defense less stable and widens the Tier~2 constraint band. The defense
cannot escape the geometry.
\end{remark}


% ============================================================
\section{From Discrete Data to Continuous Theory}
\label{sec:bridge}

\begin{theorem}[Continuous Relaxation]
\label{thm:tietze}
Let $X$ be a connected, normal, Hausdorff space. Let $S \subset X$ be a
finite set of behavioral observations with $g(p) < \tau$ for some $p$
and $g(q) > \tau$ for some $q$.

Then there exists a continuous function $f\colon X \to \R$ agreeing
with all observations for which \Cref{thm:main,thm:eps-robust} hold.
If $f$ is additionally Lipschitz (as it is for GP posterior means under
standard kernel regularity assumptions, e.g., jointly continuous kernel
on a compact domain), \Cref{thm:persistent} applies wherever the
transversality condition is met.
\end{theorem}

\begin{proof}
Finite subsets of $T_1$ spaces are closed. By the Tietze extension
theorem, $g$ extends to a continuous $f$ agreeing at every sample point.
Since $f(p) < \tau < f(q)$, \Cref{thm:main} applies. Lipschitz
extensions (which exist by McShane--Whitney) enable
\Cref{thm:eps-robust,thm:persistent}.
\end{proof}

\begin{remark}
The interpolant is not unique. All three impossibilities hold for
\emph{every} continuous extension consistent with the data. This does
not claim the results hold for the underlying discrete token system
directly, but for every continuous model consistent with the observations.
\end{remark}

\begin{corollary}
\Cref{thm:tietze} applies in $\R^n$ for every $n \geq 1$.
\end{corollary}


% ============================================================
\section{Direct Discrete Impossibility}
\label{sec:discrete}

The Tietze bridge connects discrete data to continuous theory, but a
natural objection remains: perhaps the impossibility is an artifact of
the continuous relaxation, not a property of the discrete system. To
address this, we prove a parallel set of results that work directly on
finite sets with no topology, no continuity, and no Tietze extension.
These are verified in Lean as \texttt{MoF\_12\_Discrete}.

\begin{theorem}[Discrete IVT]
\label{thm:disc-ivt}
Let $f\colon \{0,\ldots,n\!+\!1\} \to \R$ with $f(0) < \tau$ and
$f(n\!+\!1) \geq \tau$. Then there exists~$i$ with $f(i) < \tau$ and
$f(i\!+\!1) \geq \tau$.
\end{theorem}

\begin{proof}
By induction: if no consecutive crossing exists and $f(0) < \tau$,
then $f(k) < \tau$ for all~$k$, contradicting $f(n\!+\!1) \geq \tau$.
\end{proof}

This is the discrete analogue of boundary existence. The next result
is the discrete analogue of boundary fixation.

\begin{theorem}[Discrete Defense Fixation]
\label{thm:disc-defense}
If $D$ fixes all safe vertices ($f(v) < \tau \Rightarrow D(v) = v$),
the boundary crossing from \Cref{thm:disc-ivt} has its safe endpoint
fixed: $D(i) = i$.
\end{theorem}

The proof is immediate: $f(i) < \tau$, so $D(i) = i$ by
utility preservation. No continuity is needed---just the identity
constraint on safe inputs.

Where the continuous and discrete paths diverge is in what replaces
continuity as the bottleneck. In the continuous setting, the defense is
limited by its Lipschitz constant. In the discrete setting, it is
limited by its \emph{capacity}---the number of distinct output
behaviors it can realize.

\begin{theorem}[Capacity Exhaustion]
\label{thm:pigeonhole}
If there are $m$ adversarial configurations and the defense can realize
fewer than $m$ distinct behaviors, at least two configurations receive
the same treatment---so at least one is mishandled.
\end{theorem}

\begin{proof}
Pigeonhole principle, verified in Lean via
\texttt{Fintype.exists\_ne\_map\_eq\_of\_card\_lt}.
\end{proof}

\begin{theorem}[Exponential Attack Surface]
\label{thm:budget}
The attack surface grows as $2^d \geq d + 1$. For any fixed budget~$B$,
there exists $d_0$ such that $2^d > B$ for all $d \geq d_0$. No
fixed-capacity defense suffices against an exponentially growing
configuration space.
\end{theorem}

Together, these results form a \emph{discrete defense trilemma}: on any
finite ordered set where $f$ crosses~$\tau$, a utility-preserving
defense must leave the safe side of the crossing fixed, and a
bounded-capacity defense must eventually be overwhelmed by the attack
surface. The continuous and discrete paths arrive at the same conclusion
by different routes.

% ============================================================
\section{Extensions: Multi-Turn, Stochastic, and Capacity Parity}
\label{sec:extensions}

A natural objection to the preceding results is that they assume a
static, deterministic, single-turn defense. Real systems are multi-turn,
stochastic, and may have capacity comparable to the attacker. We show
that the impossibility survives all three extensions.

\subsection{Multi-Turn Impossibility}

Modern LLMs maintain state across conversation turns. One might hope
that a defense could use accumulated context to progressively eliminate
boundary vulnerabilities. This does not work.

\begin{theorem}[Multi-Turn Impossibility]
\label{thm:multi-turn}
Let $\{f_t, D_t\}_{t=1}^T$ be a sequence of alignment functions and
defenses over $T$ turns, where each $f_t$ is continuous with safe and
unsafe regions, and each $D_t$ is continuous and utility-preserving.
Then for \emph{every} turn~$t$, there exists $z_t$ with
$f_t(z_t) = \tau$ and $D_t(z_t) = z_t$.
\end{theorem}

\begin{proof}
Apply \Cref{thm:main} to $(f_t, D_t)$ at each turn. The functions
$f_t$ and $D_t$ may depend on the full history of previous turns---this
does not matter. At each timestep, the defense faces a fresh instance
of boundary fixation on the current alignment surface.
\end{proof}

The impossibility does not accumulate or diminish over turns---it
simply recurs. Memory, context, and adaptive strategies do not change
the topology of the current turn's alignment surface. But the
multi-turn setting is worse than ``the same problem repeated.'' The
attacker gains several structural advantages.

\paragraph{Information accumulation.}
At each turn, the defense's response reveals whether the probed input
is safe, unsafe, or on the boundary. Over $T$ turns with distinct
probes, the attacker classifies $T$ points, progressively narrowing the
boundary's location. The defense leaks information; the attacker
consumes it.

\paragraph{Monotone attacker improvement.}
The attacker's best observed alignment deviation score across turns is
non-decreasing: the running maximum of a sequence never drops
(verified in Lean as \texttt{running\_max\_monotone}). Each turn either
improves the attacker's best known exploit or leaves it unchanged.

\paragraph{Steering toward transversality.}
If the attacker can influence the alignment surface---for instance, by
crafting conversation context that shifts the landscape---and the
directional slope at the boundary varies continuously with the
attacker's influence parameter~$\alpha$, then the intermediate value
theorem guarantees the existence of $\alpha^*$ at which the slope
crosses the transversality threshold $L(K+1)$. For $\alpha > \alpha^*$,
Tier~3 applies: persistent unsafe volume. The attacker can find
$\alpha^*$ by binary search (verified in Lean as
\texttt{transversality\_reachable}).

\paragraph{No progressive safety.}
As long as both safe and unsafe regions exist at each turn, every turn
produces a boundary fixed point---no turn is ``free.'' The only escape
is to make $U_\tau = \emptyset$ at some turn, which requires changing
the alignment surface~$f_t$ itself (e.g., via in-context learning that
suppresses all unsafe behavior), not merely preprocessing inputs
through~$D_t$.

\subsection{Stochastic Defense Impossibility}

If the defense is randomized---mapping each input to a distribution
over outputs rather than a fixed point---one might hope that the
boundary fixation dissolves in expectation.

\begin{theorem}[Stochastic Defense Impossibility]
\label{thm:stochastic}
Let $f\colon X \to \R$ be continuous on a connected Hausdorff space
with $S_\tau, U_\tau \neq \emptyset$. Let $D$ be a stochastic defense,
and define $g(x) = \mathbb{E}_{y \sim D(x)}[f(y)]$ (the expected
alignment deviation after defense). If $g$ is continuous and $g(x) =
f(x)$ for all $x \in S_\tau$ (utility preservation in expectation),
then there exists $z$ with $f(z) = \tau$ and $g(z) = \tau$.
\end{theorem}

\begin{proof}
The function $h = g - f$ is continuous and vanishes on $S_\tau$. Since
$S_\tau$ is dense in $\cl{S_\tau}$ and $\{h = 0\}$ is closed, $h$
vanishes on $\cl{S_\tau}$. Any $z \in \cl{S_\tau} \setminus S_\tau$
satisfies $f(z) = \tau$ (by the argument of \Cref{thm:main}) and
$g(z) = f(z) = \tau$. The defense cannot reduce even the
\emph{expected} alignment deviation below $\tau$ at these points.
\end{proof}

Randomization does not help: the expected alignment deviation after a
stochastic defense inherits the same boundary fixation as a
deterministic one.

\subsection{Capacity Parity}

Even when the defense has the same computational capacity as the
attacker, utility preservation acts as a structural tax.

\begin{theorem}[Capacity Parity Disadvantage]
\label{thm:capacity}
Let the defense and attacker both operate on a space of $N$ inputs,
of which $|S_\tau|$ are safe. The defense must implement identity on
all safe inputs, consuming $|S_\tau|$ of its capacity. The attacker
faces no such constraint. Under capacity parity ($C = N$), the
defense's effective free capacity is $N - |S_\tau| < N$, giving the
attacker a permanent edge of $|S_\tau|$ capacity units.
\end{theorem}

This is a counting argument: utility preservation is not free. Every
safe input the defense must leave untouched is one fewer degree of
freedom available for remapping unsafe inputs. The attacker, who has
no obligation to preserve anything, retains full capacity.


% ============================================================
\section{The Vulnerability Landscape}
\label{sec:landscape}

\begin{theorem}[Basin Structure]
\label{thm:basin}
If $f$ is continuous and $f(p)>\tau$, then $U_\tau$ is open. Under
any measure that assigns positive mass to nonempty open sets (e.g.,
Lebesgue measure on~$\R^n$), $U_\tau$ has positive measure.
\end{theorem}

\begin{theorem}[Basin Fragment Minimum Size]
\label{thm:fragment}
If $f$ is $L$-Lipschitz with $f(p) > \tau$, the connected component
of $U_\tau$ containing $p$ has diameter $\geq 2(f(p) - \tau)/L$.
\end{theorem}


% ============================================================
\section{Attack Properties}
\label{sec:attacks}

\begin{theorem}[Perturbation Robustness]
\label{thm:lipschitz}
If $f$ is $L$-Lipschitz and $f(p)>\tau$, then
$\operatorname{ball}(p,\, (f(p)-\tau)/L) \subseteq U_\tau$.
The radius is monotone in~$f(p)$.
\end{theorem}

\begin{theorem}[Iterative Convergence]
\label{thm:convergence}
Any monotone-improvement operator $T$ with score in $[0,1]$ converges.
If each step gains $\geq \delta$, convergence takes
$\leq \lfloor 1/\delta \rfloor$ steps.
\end{theorem}

\begin{theorem}[Transferability]
\label{thm:transfer}
$\|f-g\|_\infty \leq \delta \implies
\{f > \tau+\delta\} \subseteq \{g > \tau\}$.
Transfer costs zero additional queries.
\end{theorem}

\begin{theorem}[Authority Monotonicity]
\label{thm:authority}
If $f(y,\cdot)$ is monotone non-decreasing in authority for each fixed
indirection~$y$, then the vulnerability set $\{x : f(y,x) > \tau\}$ is
upward-closed and a critical threshold $a^*_2(y)$ exists (by IVT, when
$f(y,\cdot)$ is continuous and crosses~$\tau$). If additionally $f$ is
monotone non-decreasing in indirection for each fixed authority, the
threshold curve $y \mapsto a^*_2(y)$ is non-increasing.
\end{theorem}

\begin{theorem}[Gradient Ascent]
\label{thm:gradient}
If $f$ has nonzero Fr\'{e}chet derivative at~$x$, there exist $v$ and
$\varepsilon > 0$ with $f(x + \varepsilon v) > f(x)$.
\end{theorem}


% ============================================================
\section{Stability Under Fine-Tuning}
\label{sec:stability}

\begin{theorem}[Interior Stability]
\label{thm:interior-stable}
If $\|f - g\|_\infty \leq \varepsilon$:
$f(x) > \tau + \varepsilon \implies g(x) > \tau$, and
$f(x) < \tau - \varepsilon \implies g(x) < \tau$.
Only the band $|f(x) - \tau| \leq \varepsilon$ is uncertain.
\end{theorem}

\begin{theorem}[Crossing Preservation]
\label{thm:crossing}
If $f$ crosses $\tau$ transversally on $[a,b]$ and $g$ is continuous
with $\|f - g\|_\infty \leq \varepsilon$, then $g$ also crosses $\tau$
on $[a,b]$.
\end{theorem}

\begin{theorem}[Patching Is Nonlocal]
\label{thm:nonlocal}
There exist $f, g$ with $\|f-g\|_\infty \leq \varepsilon$ such that
eliminating a vulnerability at one point necessarily changes values at
distant points. (Verified by explicit construction in Lean.)
\end{theorem}


% ============================================================
\section{Cost Asymmetry}
\label{sec:cost}

\begin{theorem}[Exponential Cost Asymmetry for Grid-Based Defense]
\label{thm:cost}
For a defense monitoring the behavioral space via a grid with
$N \geq 2$ bins per axis in $d$ dimensions: attack cost
$\leq 1/\delta$ (dimension-independent); grid-based defense cost
$= N^d$ (exponential in~$d$); the ratio
$\delta \cdot N^d \to \infty$ as $d\to\infty$.
\end{theorem}

At $d=2$ with $N=25$ and $\delta=0.01$, the ratio is $6.25$.
At $d=10$, it climbs to $\sim 10^{12}$. Verified by \texttt{norm\_num}
in Lean.


% ============================================================
\section{Experimental Validation}
\label{sec:experiments}

We validate against the Manifold of Failure
framework~\cite{bhatt2025manifold}, which maps three LLMs over a 2D
behavioral space (query indirection $\times$ authority framing).

\begin{table}[t]
\centering
\small
\caption{Falsifiable predictions confirmed by empirical data.}
\label{tab:correspondence}
\begin{tabular}{@{}p{3.0cm}p{4.5cm}p{4.5cm}@{}}
\toprule
\textbf{Theorem} & \textbf{Predicts} & \textbf{Confirmed by} \\
\midrule
Basin Structure (\ref{thm:basin})
  & Basins are open with positive measure
  & Heatmaps show extended regions \\[3pt]
Fragmentation (\ref{thm:fragment})
  & Smooth $\to$ large basins; rough $\to$ mosaic
  & Llama: mesa; GPT-OSS: mosaic \\[3pt]
Convergence (\ref{thm:convergence})
  & Attacks exhibit monotone convergence
  & Convergence curves plateau \\[3pt]
Transferability (\ref{thm:transfer})
  & Similar surfaces $\to$ shared basins
  & Llama $.93 \to$ GPT-OSS $.73 \to$ Mini $.47$ \\[3pt]
Authority (\ref{thm:authority})
  & Horizontal banding
  & Bands at $a_2 \approx .25$--$.35$, $.65$--$.85$ \\[3pt]
Persistent (\ref{thm:persistent})
  & Steep boundaries $\to$ unsafe volume persists
  & Llama's $.93$ plateau immune to defense \\[3pt]
Interior Stability (\ref{thm:interior-stable})
  & Deep basin points survive fine-tuning
  & Vulnerability persists across variants \\[3pt]
Cost (\ref{thm:cost})
  & 2D tractable, high-$d$ intractable
  & 15K queries fill 63\% at $d\!=\!2$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Llama-3-8B} (mean AD $0.93$, basin rate $93.9\%$): near-flat
mesa, small~$L$, large robustness radii. The directional slope of the
alignment surface at boundary crossings far exceeds $L(K+1)$ for any
reasonable~$K$, placing this model squarely in Tier~3 territory: the
defense provably cannot eliminate the vulnerability volume.

\paragraph{GPT-OSS-20B} (mean AD $0.73$, basin rate $64.3\%$): rugged
landscape, large~$L$, small fragments. Horizontal bands confirm
authority monotonicity. The fragmented boundary presents many
transversal crossings, each generating a persistent unsafe pocket.

\paragraph{GPT-5-Mini} (peak AD $0.50$, basin rate $0\%$): at
$\tau = 0.50$, $U_\tau = \emptyset$, so none of the three tiers
apply---correctly predicting no impossibility. The model's strategy:
make the boundary harmless rather than eliminate it.

\paragraph{Quantitative checks.}
Robustness radius: $L \approx 5$ for Llama gives
$r = (0.93-0.50)/5 = 0.086$, covering $\sim$2 grid cells---consistent
with heatmap smoothness. Cost ratio: $0.01 \times 625 = 6.25$
theoretical vs.\ ${\sim}4$ empirical.


% ============================================================
\section{The Engineering Prescription}
\label{sec:engineering}

The impossibility results do not say that defense is futile---they say
that \emph{perfect} defense is impossible under the stated constraints.
The gap between impossibility and practice is where engineering lives.
The following prescriptions are ordered from most to least actionable.

\paragraph{1. Make the boundary shallow.}
Set $\tau$ so that boundary-level behavior is benign. If $f(z) = \tau$
corresponds to a polite refusal rather than harmful compliance, the
impossibility is mathematically true but practically harmless.
GPT-5-Mini exemplifies this: its ceiling at $\AD = 0.50$ means the
``Manifold of Failure'' exists but contains no actual harm.

\paragraph{2. Reduce the Lipschitz constant.}
Smaller~$L$ makes the alignment surface smoother and more predictable.
It also raises the transversality threshold $L(K+1)$ for Tier~3: a
smoother surface is harder for the attacker to steer into persistent
unsafe territory. The tradeoff is that smoother surfaces spread
vulnerabilities over wider regions, but those regions are more easily
monitored.

\paragraph{3. Reduce the effective dimension.}
Defense cost grows as $N^d$. Constraining the prompt
interface---standardized formats, restricted API parameters, bounded
context lengths---reduces~$d$, making the behavioral space tractable.

\paragraph{4. Monitor, don't eliminate, the boundary.}
\Cref{thm:crossing} shows transversal crossings persist under
fine-tuning, and \Cref{thm:multi-turn} shows they recur every turn.
Rather than attempting the impossible (eliminating boundary points),
deploy runtime monitoring that detects approach to the boundary. The
Lipschitz bound (\Cref{thm:lipschitz}) gives a computable estimate of
distance to the boundary from any observed AD value.

\paragraph{5. Abandon stateful security illusions.}
\Cref{thm:multi-turn} proves that accumulating context across turns
does not eliminate boundary fixation---the impossibility recurs fresh
at every turn. Defenses that rely on ``learning'' from conversation
history to progressively close vulnerabilities are fighting topology,
not winning. Each turn is a new instance of the same problem.

\paragraph{6. Do not pay the stochastic tax.}
\Cref{thm:stochastic} proves that randomized defenses---stochastic
routing, temperature-based sampling, random prompt perturbation---fail
in expectation. The expected alignment deviation after a stochastic
defense inherits the same boundary fixation as a deterministic one.
Compute spent on randomization is compute wasted, at least for the
purpose of eliminating boundary points.

\paragraph{7. Break capacity parity with architectural separation.}
\Cref{thm:capacity} proves that under equal resources, the defense
loses capacity to utility preservation while the attacker retains full
freedom. The implication: a defense that shares the same architecture
and parameter budget as the model it protects starts at a structural
disadvantage. To break this asymmetry, separate the defense from the
utility pipeline entirely---for example, by using a dedicated
lightweight classifier that does not need to preserve utility (it only
flags), paired with a separate response generator that handles safe
inputs. This sacrifices the ``single continuous wrapper'' model but
sidesteps the capacity tax.


% ============================================================
\section{Why the Impossibility Is Inescapable}
\label{sec:meta}

The preceding sections establish impossibility via multiple routes:
continuous topology, Lipschitz metric structure, discrete counting,
multi-turn dynamics, stochastic expectations, and capacity constraints.
It is natural to ask whether these are separate results that happen to
agree, or instances of a single underlying phenomenon.

They are instances of a single phenomenon.

\begin{theorem}[Representation-Independent Meta-Theorem]
\label{thm:meta}
Let $X$ be any set, $S \subseteq X$ nonempty and proper (the safe
region), and $D\colon X \to X$ a defense satisfying:
\begin{enumerate}[nosep]
  \item \emph{Utility preservation}: $D(x) = x$ for all $x \in S$.
  \item \emph{Bounded regularity}: the fixed-point set
    $\Fix(D) = \{x : D(x) = x\}$ strictly contains $S$.
\end{enumerate}
Then there exists $x \notin S$ with $D(x) = x$---the defense fixes a
non-safe input.
\end{theorem}

\begin{proof}
$\Fix(D) \supsetneq S$ implies $\Fix(D) \setminus S \neq \emptyset$.
Any $x$ in this difference satisfies $x \notin S$ and $D(x) = x$.
\end{proof}

The content is not in the conclusion---which is a triviality of set
theory---but in the \emph{regularity condition} that forces
$\Fix(D) \supsetneq S$. Each path through the theory provides a
different proof that this condition holds:

\begin{center}
\small
\begin{tabular}{@{}lp{8cm}@{}}
\toprule
\textbf{Path} & \textbf{Why $\Fix(D) \supsetneq S$} \\
\midrule
Continuous (Tier~1)
  & $\Fix(D)$ is closed (Hausdorff), $S$ is open and not closed
    (connected), so $\cl{S} \subseteq \Fix(D)$ but
    $\cl{S} \supsetneq S$. \\[3pt]
Lipschitz (Tier~2)
  & Same as Tier~1, plus Lipschitz bounds constrain how far the
    defense can move points near the fixed boundary. \\[3pt]
Transversal (Tier~3)
  & Same as Tier~2, plus steep slope forces $f(D(x)) > \tau$
    on positive-measure set. \\[3pt]
Discrete
  & IVT on finite ordered set forces a crossing; utility preservation
    fixes the safe endpoint. \\[3pt]
Stochastic
  & $g - f$ vanishes on $S$ and is continuous, so it vanishes on
    $\cl{S}$; expected AD $= \tau$ at boundary. \\[3pt]
Multi-turn
  & Regularity holds at \emph{each} turn independently. \\[3pt]
Capacity
  & Finite capacity $<$ attack surface forces pigeonhole collision. \\
\bottomrule
\end{tabular}
\end{center}

\noindent For $\Fix(D)$ to equal $S$ exactly---the only way to escape
the meta-theorem---$S$ would need to be simultaneously open (as a
preimage of an open set under a continuous $f$) and closed (as a
fixed-point set of a continuous $D$). In a connected space with both
safe and unsafe inputs, no proper nonempty subset is both open and
closed. This is the geometric lock.

The meta-theorem is representation-independent: it does not care whether
$X$ is $\R^n$, a discrete token space, a probability simplex, or an
abstract set. It requires only that some mechanism---topological,
metric, combinatorial, or information-theoretic---forces the
fixed-point set to spill beyond the safe region.


% ============================================================
\section{The Lean Artifact}
\label{sec:artifact}

The complete theory is verified in Lean~4.28.0 with Mathlib~v4.28.0.
The artifact comprises 38 files:
\begin{itemize}[nosep]
  \item 10 core theory files (MoF\_01--MoF\_10)
  \item 10 cost theory files (MoF\_Cost\_01--MoF\_Cost\_10)
  \item 10 advanced theory files (MoF\_Adv\_01--MoF\_Adv\_10)
  \item 1 continuous relaxation (MoF\_ContinuousRelaxation)
  \item 1 $\varepsilon$-robust + persistent unsafe region
    (MoF\_11\_EpsilonRobust)
  \item 1 discrete impossibility (MoF\_12\_Discrete)
  \item 1 multi-turn + stochastic + capacity parity
    (MoF\_13\_MultiTurn)
  \item 1 representation-independent meta-theorem
    (MoF\_14\_MetaTheorem)
  \item 3 capstone files (MasterTheorem, Euclidean instantiation,
    verification)
\end{itemize}

Every theorem depends on exactly three axioms: \texttt{propext},
\texttt{Classical.choice}, \texttt{Quot.sound}---Lean's standard
foundations. No additional axioms.

Key Mathlib lemmas used:
\begin{center}
\small
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Lean name} & \textbf{Role} \\
\midrule
\texttt{isClosed\_diagonal} & Hausdorff $\Rightarrow$ diagonal closed \\
\texttt{LipschitzWith.dist\_le\_mul} & Lipschitz distance bound \\
\texttt{isClopen\_iff} & Connected $\Rightarrow$ no proper clopen \\
\texttt{closure\_minimal} & Closure in closed superset \\
\texttt{IsOpen.measure\_pos} & Open nonempty $\Rightarrow$ positive measure \\
\texttt{intermediate\_value$_2$} & IVT on connected spaces \\
\bottomrule
\end{tabular}
\end{center}


% ============================================================
\section{Limitations}
\label{sec:limitations}

\paragraph{Tier~1 is at the boundary.}
Fixed points satisfy $f(z) = \tau$ exactly. If $\tau$-level behavior
is benign, the theorem is true but harmless.

\paragraph{Tier~2 constrains depth, not direction.}
The defense may push near-boundary points slightly below~$\tau$. The
bound limits how far, not whether.

\paragraph{Tier~3 requires transversality.}
The persistent unsafe region exists only where the alignment surface
is steep ($c > L(K+1)$). Flat or gently sloping boundaries escape
Tier~3 (though not Tiers~1--2).

\paragraph{The Tietze bridge applies to interpolants.}
The impossibilities hold for every continuous extension of discrete
data, not for the underlying discrete system itself.

\paragraph{Grid-based cost asymmetry.}
\Cref{thm:cost} assumes exhaustive grid enumeration. Learning-based
defenses that generalize across the space may sidestep the exponential
bound.

\paragraph{Single-turn only.}
The manifold may shift across conversation turns.


% ============================================================
\section{Conclusion}
\label{sec:conclusion}

We have established three impossibility results for prompt injection
defense, forming a progressive hierarchy. Tier~1 (boundary fixation):
any continuous, utility-preserving defense must fix boundary points.
Tier~2 ($\varepsilon$-robust constraint): the defense cannot uniformly
reduce alignment deviation far below~$\tau$ on a positive-measure band.
Tier~3 (persistent unsafe region): under transversality, the defense
leaves positive-measure volume genuinely unsafe---$f(D(x)) > \tau$
persists.

A representation-independent meta-theorem (\Cref{thm:meta}) reveals the
common structure: every path through the theory proves the same thing---that
the defense's fixed-point set must spill beyond the safe region.
Continuous topology, Lipschitz bounds, discrete counting, stochastic
expectations, multi-turn dynamics, and capacity constraints all force
this spillover through different mechanisms, but the conclusion is
identical. The impossibility is not an artifact of any particular
representation; it is a consequence of combining utility preservation
with any form of bounded regularity.
Supporting theorems show that vulnerabilities come in
basins, attacks converge and transfer, vulnerabilities survive
fine-tuning, and defense cost grows exponentially. Each result is
validated against empirical data from three frontier LLMs.

The practical prescription: make the boundary shallow (raise~$\tau$),
smooth (reduce~$L$), and low-dimensional (constrain the interface).
GPT-5-Mini's ceiling at $\AD = 0.50$ shows this working in practice.

The full theory is machine-verified in Lean~4: 38 files, approximately
300 theorems, zero \texttt{sorry}, three standard axioms.


% ============================================================
\bibliographystyle{plain}
\begin{thebibliography}{99}

\bibitem{alon2023detecting}
G.~Alon and M.~Kamfonas.
\newblock Detecting language model attacks with perplexity.
\newblock \emph{arXiv preprint arXiv:2308.14132}, 2023.

\bibitem{bagnall2019certifying}
A.~Bagnall and G.~Stewart.
\newblock Certifying the true error: Machine learning in {Coq} with
  verified generalization guarantees.
\newblock \emph{Proceedings of the AAAI Conference on Artificial
Intelligence}, 2019.

\bibitem{bai2022constitutional}
Y.~Bai et~al.
\newblock Constitutional {AI}: Harmlessness from {AI} feedback.
\newblock \emph{arXiv preprint arXiv:2212.08073}, 2022.

\bibitem{bhatt2025manifold}
M.~Bhatt, S.~Munshi, V.~S.~Narajala, I.~Habler, A.~Al-Kahfah,
K.~Huang, and B.~Gatto.
\newblock Manifold of failure: Behavioral attraction basins in language
  models.
\newblock \emph{arXiv preprint arXiv:2602.22291v2}, 2026.

\bibitem{bolte2024}
J.~Bolte et~al.
\newblock Formal verification of optimization algorithms.
\newblock \emph{Mathematical Programming}, 2024.

\bibitem{chao2024jailbreaking}
P.~Chao, A.~Robey, E.~Dobriban, H.~Hassani, G.~J. Pappas, and
E.~Wong.
\newblock Jailbreaking black box large language models in twenty queries.
\newblock \emph{arXiv preprint arXiv:2310.08419}, 2024.

\bibitem{cohen2019certified}
J.~Cohen, E.~Rosenfeld, and J.~Z. Kolter.
\newblock Certified adversarial robustness via randomized smoothing.
\newblock \emph{Proceedings of ICML}, 2019.

\bibitem{goodfellow2014explaining}
I.~J. Goodfellow, J.~Shlens, and C.~Szegedy.
\newblock Explaining and harnessing adversarial examples.
\newblock \emph{arXiv preprint arXiv:1412.6572}, 2014.

\bibitem{inan2023llama}
H.~Inan et~al.
\newblock {Llama Guard}: {LLM}-based input-output safeguard for
  human-{AI} conversations.
\newblock \emph{arXiv preprint arXiv:2312.06674}, 2023.

\bibitem{katz2017reluplex}
G.~Katz, C.~Barrett, D.~L. Dill, K.~Julian, and
M.~J. Kochenderfer.
\newblock Reluplex: An efficient {SMT} solver for verifying deep neural
  networks.
\newblock \emph{Proceedings of CAV}, 2017.

\bibitem{mehrotra2024tree}
A.~Mehrotra et~al.
\newblock Tree of attacks: Jailbreaking black-box {LLMs} automatically.
\newblock \emph{arXiv preprint arXiv:2312.02119}, 2024.

\bibitem{mouret2015illuminating}
J.-B. Mouret and J.~Clune.
\newblock Illuminating search spaces by mapping elites.
\newblock \emph{arXiv preprint arXiv:1504.04909}, 2015.

\bibitem{samvelyan2024rainbow}
M.~Samvelyan et~al.
\newblock Rainbow teaming: Open-ended generation of diverse adversarial
  prompts.
\newblock \emph{Advances in Neural Information Processing Systems}, 37,
2024.

\bibitem{singh2019abstract}
G.~Singh, T.~Gehr, M.~P{\"u}schel, and M.~Vechev.
\newblock An abstract domain for certifying neural networks.
\newblock \emph{Proceedings of POPL}, 2019.

\bibitem{szegedy2013intriguing}
C.~Szegedy et~al.
\newblock Intriguing properties of neural networks.
\newblock \emph{arXiv preprint arXiv:1312.6199}, 2013.

\bibitem{tsipras2018robustness}
D.~Tsipras, S.~Santurkar, L.~Engstrom, A.~Turner, and A.~Madry.
\newblock Robustness may be at odds with accuracy.
\newblock \emph{arXiv preprint arXiv:1805.12152}, 2018.

\bibitem{wolpert1997no}
D.~H. Wolpert and W.~G. Macready.
\newblock No free lunch theorems for optimization.
\newblock \emph{IEEE Trans.\ Evol.\ Comput.}, 1(1):67--82, 1997.

\bibitem{zou2023universal}
A.~Zou, Z.~Wang, N.~Carlini, M.~Nasr, J.~Z. Kolter, and
M.~Fredrikson.
\newblock Universal and transferable adversarial attacks on aligned
  language models.
\newblock \emph{arXiv preprint arXiv:2307.15043}, 2023.

\end{thebibliography}

\end{document}
